---
title: "Exercise 2"
output:
  html_document:
    df_print: paged
  pdf_document: default
date: "2024-01-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Group members:

* Tashfeen Ahmed
* Adrian Alarcon
* Yvan Kammelu
* Zhicheng Zhong


```{r}
#install.packages(c("arrow","gender", "wru", "lubridate", "gtsummary"))
# Load required libraries
library(gender)
library(wru)
library(lubridate)
library(dplyr)
library(gtsummary)
library(arrow)
library(tidyr)
library(zoo)
library(purrr)


```

```{r}
data<- read_feather("app_data_starter.feather")
```



```{r}
# Task 1: Create individual-level variables
examiner_names <- data %>% distinct(examiner_name_first)

examiner_names
```

## Obtaining gender of the examiner

Using the `gender` package, we identify the gender of the examiner based on the first name, according to the documentation.

```{r}
# get a table of names and gender
data$examiner_name_first

examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )


examiner_names_gender
```
In this part, we joined the gender data obtained in the previous step into the main dataset.

```{r}
# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)

# joining gender back to the dataset
data <- data %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")

# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()
```
## Obtaining the race of the examiner

Based on the last name, and using the `wru` package, we identified the probability of the examiner to be of an specific race among Asian, Black, Hispanic and other.

```{r}
library(wru)

examiner_surnames <- data %>% 
  select(surname = examiner_name_last) %>% 
  distinct()

examiner_surnames
```



```{r}
examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()
examiner_race
```

```{r}
examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))

examiner_race
```

On this step, we cleaned the dataset removing extra columns

```{r}
# removing extra columns
examiner_race <- examiner_race %>% 
  select(surname,race)

data <- data %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))

rm(examiner_race)
rm(examiner_surnames)
gc()
```


```{r}
library(lubridate) # to work with dates

examiner_dates <- data %>% 
  select(examiner_id, filing_date, appl_status_date) 

examiner_dates
```


```{r}
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))
```

After the cleaning and preprocessing steps, we grouped the data at a examiner level. This would allow us to perform a regression models

```{r}
examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
    ) %>% 
  filter(year(latest_date)<2018)

examiner_dates
```

```{r}
data <- data %>% 
  left_join(examiner_dates, by = "examiner_id")

rm(examiner_dates)
gc()
```
```{r}
data
data <- data %>%
  select(
    application_number,
    filing_date,
    examiner_name_last,
    examiner_name_first,
    examiner_name_middle,
    examiner_id,
    examiner_art_unit,
    uspc_class,
    uspc_subclass,
    patent_number,
    patent_issue_date,
    abandon_date,
    disposal_type,
    appl_status_code,
    appl_status_date,
    tc,
    gender = gender.y,  # Renaming the column to remove the suffix
    race = race.y,      # Renaming the column to remove the suffix
    earliest_date = earliest_date.y,  # Renaming the column to remove the suffix
    latest_date = latest_date.y,  # Renaming the column to remove the suffix
    tenure_days = tenure_days.y   # Renaming the column to remove the suffix
  )
data
```


# Task 2: Create a panel dataset
# ------------------------------

```{r}
library(dplyr)
library(lubridate)
library(zoo)

# Convert dates to quarters
data <- data %>%
  mutate(
    filing_year_quarter = as.yearqtr(filing_date),
    abandon_year_quarter = as.yearqtr(abandon_date),
    issue_year_quarter = as.yearqtr(patent_issue_date)
  )

# Aggregate applications data by quarter
panel_data <- data %>%
  group_by(examiner_id, filing_year_quarter) %>%
  summarise(
    num_new_applications = n_distinct(application_number),
    num_abandoned_applications = sum(disposal_type == "ABN", na.rm = TRUE),
    num_issued_patents = sum(disposal_type == "ISS", na.rm = TRUE),
    num_in_process_applications = sum(disposal_type == "PEND", na.rm = TRUE),
    current_art_unit = first(examiner_art_unit),
    .groups = 'drop'
  )

# Add the count of people and women in each art unit per quarter
art_unit_info <- data %>%
  group_by(filing_year_quarter, examiner_art_unit) %>%
  summarise(
    num_people_in_art_unit = n_distinct(examiner_id),
    num_women_in_art_unit = sum(gender == "female", na.rm = TRUE),
    .groups = 'drop'
  )

# Join the art unit info with the main panel data
panel_data <- panel_data %>%
  left_join(art_unit_info, by = c("filing_year_quarter", "current_art_unit" = "examiner_art_unit"))

# Mark the last five quarters for each examiner
panel_data <- panel_data %>%
  group_by(examiner_id) %>%
  mutate(
    # Get a list of the last five quarters of activity for each examiner
    last_five_quarters = list(tail(sort(unique(filing_year_quarter)), 5))
  ) %>%
  ungroup() %>%
  mutate(
    # Check if the current quarter is in the last five quarters of activity
    separation_indicator = if_else(map_lgl(filing_year_quarter, ~ .x %in% last_five_quarters[[1]]), 1, 0)
  )


# Detect changes in current_art_unit
panel_data <- panel_data %>%
  group_by(examiner_id) %>%
  mutate(
    # If the current art unit is different from the previous one, it's a move (1), otherwise, it's not (0).
    # For the first row of each examiner (where there is no "previous" art unit), use NA as the default value.
    AU_move_indicator = if_else(current_art_unit != lag(current_art_unit, default = NA), 1, 0)
  ) %>%
  mutate(
    # Replace NA with 0 - assumes that the first observation is not a move.
    AU_move_indicator = replace_na(AU_move_indicator, 0)
  ) %>%
  ungroup()




```



```{r}
table(panel_data$separation_indicator)
table(panel_data$AU_move_indicator)

```


# Task 3: Estimate predictors for turnover and mobility
# ---------------------------------------------------


```{r}
# Prepare the data for regression
regression_data <- panel_data %>%
  filter(num_new_applications > 0)

# Regression model for Turnover
turnover_model <- glm(separation_indicator ~  num_new_applications + num_abandoned_applications + 
                      num_issued_patents +  
                      num_people_in_art_unit + num_women_in_art_unit,
                    family = binomial(), data = regression_data)

# Regression model for Mobility (AU Move)
mobility_model <- glm(AU_move_indicator ~ num_new_applications + num_abandoned_applications +
                      num_issued_patents + num_in_process_applications + 
                      num_people_in_art_unit + num_women_in_art_unit,
                    family = binomial(), data = regression_data)

# Create descriptive tables for both models
turnover_table_ <- tbl_regression(turnover_model)

turnover_table_
```

Showing the models

```{r}
turnover_model
```


```{r}
mobility_mode_ <- tbl_regression(mobility_model)
print(mobility_mode_)
```


```{r}
mobility_model
```




